---
title: "Assignment 10"
author: "Mihir Parikh"
date: "3/20/2018"
output:
  html_document:
    keep_md: true
    fig_width: 7
    fig_height: 7
---

```{r libraries, include=TRUE, message=FALSE, warning=FALSE}
library("rvest")
library("dplyr")
library("data.table")
library("ggplot2")
```

<center><h1>Question 1: Mental Health Clinics</h1></center>

### a. Load Data
```{r load_data}
load("N-MHSS-2015-DS0001-data-r.rda")
mh2015_puf <- data.frame(mh2015_puf)
```

### b. State Abbreviations
```{r state_abbreviations}
levels(unique(mh2015_puf$LST))
```

### c. Mainland Locations
```{r mainland_locations}
omitted <- c('AK','HI','VI','PR','PW','MP','MH','GU','FM','AS')
mainlandStates <- mh2015_puf %>% filter(!trimws(as.character(LST)) %in% omitted)
stateSummary <- mainlandStates %>% group_by(LST) %>% summarise(FREQUENCY = n()) %>% data.frame()
colnames(stateSummary)[1] <- "STATE"
stateSummary
```

### d. Frequency Plot
```{r frequency_plot}
ggplot(data=stateSummary, aes(reorder(STATE, FREQUENCY),FREQUENCY)) +
  geom_bar(stat="identity", aes(fill=STATE))+ 
  ggtitle('Frequency of VA Medical Centers by State in Mainland USA') + coord_flip() +
  xlab('Mainland States of VA Medical Centers ') +
  ylab('Frequency') +
  theme(plot.title=element_text(hjust = .5)) +
  theme(axis.text.y = element_text(hjust=2)) + theme(legend.position="none")
```

<center><h1>Question 2: Cleaning and Bringing In New Features</h1></center>

### a. Try to Merge
```{r attempted_merge}
# The problem is that there are extra spaces in the State abbreviation names in the stateSummary dataframe. This causes the frequencies to be null.

statesize <- read.csv(file="statesize.csv", header=TRUE, sep=",")
colnames(stateSummary)[1] <- "Abbrev"
merge(stateSummary, statesize, by="Abbrev", all.y=TRUE)
```

### b. Fix and Merge
```{r merge}
stateSummary$Abbrev = gsub(' ','',stateSummary$Abbrev)
merged <- merge(stateSummary, statesize, by="Abbrev")
merged
```

### c. Add Hospitals Per Thousand Square Miles
```{r add_hospitals_per_thousand_sq_miles}
merged$MedCentersPerThsdSqMiles <- merged$FREQUENCY/(merged$SqMiles/1000)
head(merged, 10)
```

### d. Frequency Plot for Hospitals Per Thousand Square Miles
```{r frequency_plot_thousand_sq_miles}
ggplot(data = merged, aes(reorder(Abbrev, MedCentersPerThsdSqMiles), MedCentersPerThsdSqMiles)) +
  geom_bar(stat="identity", aes(fill = Region)) + 
  ggtitle('Density of VA Medical Centers Per Thousand Square Miles in Mainland USA') + coord_flip() +
  scale_fill_brewer(palette="Dark2") +
  xlab('Mainland States') +
  ylab('Density per State (Medical Centers Per 1000 Square Miles)') +
  theme(plot.title=element_text(hjust = .5)) +
  theme(axis.text.y = element_text(hjust=1))
```

### e. What patterns are seen?

Smaller states have higher frequencies of medical centers per thousand miles while larger states tend to have smaller frequencies. The Northeast appears to have a high frequency while the West appears to have the lowest frequency. I would request to my boss that we address looking at population density and creating an aggregate metric that weights the frequency based on distance with the frequency based on population.

Github link: https://github.com/mihir787/msds_homework/tree/master/assignment10_exploratory_data